koda: Deep Learning-Enhanced Pipeline for Keyword Detection in Document Images

### 2020-01-10 paper/paper.tex
First draft of the paper

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage[margin=3cm]{geometry}

\graphicspath{{images/}}

@@

The paper is organized as follows: the first section, Architecture, describes the general structure and components of KODA. In the Edge Detection section, starting from the limitations of traditional techniques, a deep learning-based approach is proposed, along with a throughout comparison. Thereafter, the Corners Detection section describes the document detection algorithm itself, exploiting Hough transformations. Then, in the Document analysis section, warping, OCR and highlighting techniques are described. Finally, the concluding remarks are drawn in the last section.

@@

\begin{figure}[H]
	\includegraphics[width=\linewidth]{Pipeline.png}
	\caption{KODA Pipeline scheme with sub-tasks. The computationally expensive processing is done once per input image, then the user can search for keywords and retrieves the output image multiple times repeatedly.}
	\label{fig:pipeline}
\end{figure}

\begin{figure}[H]
	\includegraphics[width=\linewidth]{ipt.png}
	\caption{Image processing tasks example. Note: edges detection returns a resized (256x256)}
	\label{fig:imageprocessing}
\end{figure}

@@

Our initial attempts to detect document edges involved the well known \textit{Canny edge detector} [TODO REF]. Although applying the filter directly does detect the sheet edges, it presents many artifacts
due to the text printed on the document itself (as seen in Fig. \ref{fig:canny_gaussian} (B)).

In order to mitigate the problem, a \textit{Gaussian filter} [TODO REF] is applied before the edge detection, obtaining a clear highlighting of the interesting edges (as seen in Fig.  \ref{fig:canny_gaussian} (C)).

@@

For this particular task, \textit{Keras}, an open-source Python library for Deep Learning [TODO REF], was chosen to build the model. Many experiments were made to find the right architecture, most of them 
exploiting \textit{convolutional neural networks}. Above all of them, KODA uses \textit{U-Net}[TODO REF], a popular architecture for \textit{object segmentation}, and in particular an open-source implementation for Keras [TODO REF].

@@

Due to the high number of required samples, the dataset itself is insufficient to fully train the network.
For this reason, a technique called \textit{data augmentation}[TODO REF] is applied to provide the model
with enough variability. In particular, before being fed to the network, a random combination of transformations, such as scale, translation, rotation and brightness changes, is applied to the image and to the resulting edge feature map, as shown in Fig. \ref{fig:augmented_image}.

@@

\begin{figure}[htb!]
	\centering
	\includegraphics[width=\linewidth]{edge_comparison_epoch.png}
	\caption{Comparison of the edge detector over different epochs.}
	\label{fig:edge_comparison_epoch}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{deep_edge_comparison.png}}%

@@

\section{Corners detection}

The corners detection task goal is to find the four document corners starting from the one-channel edges image. These points should be as precise as possible in order to define the tightest convex polygon around the document. The intuitive method to solve this task is to use the Hough transformation, as suggested in a Dropbox article \footnote{https://blogs.dropbox.com/tech/2016/08/fast-and-accurate-document-detection-for-scanning/}, to find straight lines that best approximate the document edges; the corners will be the intersection between those lines.

@@

To maintain uniformity for the client APi, the corners are sorted following the Cartesian quadrants and they are scaled up based on the original image size provided as input.

@@

In order to implement the \textit{Optical Character Recognition} feature needed to highlight keywords, KODA uses \textit{Tesseract}[TODO REF], an open-source OCR engine by Google. Moreover, it uses \textit{TesserOCR}[TODO REF], a Python wrapper which provides a fast binding to the C++ native library and exposes all the low-level APIs.

>>>>

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage[margin=3cm]{geometry}
\usepackage{titlesec}

\graphicspath{{images/}}
\newcommand{\sectionbreak}{\clearpage}

@@

The paper is organized as follows: the first section, Architecture, describes the general structure and components of KODA. In the Edge Detection section, starting from the limitations of traditional techniques, a deep learning-based approach is proposed, along with a thorough comparison. Thereafter, the Corners Detection section describes the document detection algorithm itself, exploiting Hough transformations. Then, in the Document analysis section, warping, OCR and highlighting techniques are described. Finally, the concluding remarks are drawn in the last section.

@@

\begin{figure}[!htbp]
	\includegraphics[width=\linewidth]{ipt.png}
	\caption{Image processing tasks example. Note: edges detection returns a resized (256x256)}
	\label{fig:imageprocessing}
\end{figure}

\begin{figure}[!htbp]
	\includegraphics[width=\linewidth]{Pipeline.png}
	\caption{KODA Pipeline scheme with sub-tasks. The computationally expensive processing is done once per input image, then the user can search for keywords and retrieves the output image multiple times repeatedly.}
	\label{fig:pipeline}
\end{figure}

@@

Our initial attempts to detect document edges involved the well known \textit{Canny edge detector} \footnote{J. Canny, "A Computational Approach to Edge Detection," in IEEE Transactions on Pattern Analysis and Machine Intelligence, vol. PAMI-8, no. 6, pp. 679-698, Nov. 1986.
	doi: 10.1109/TPAMI.1986.4767851}. 
Although applying the filter directly does detect the sheet edges, it presents many artifacts
due to the text printed on the document itself (as seen in Fig. \ref{fig:canny_gaussian} (B)).

In order to mitigate the problem, a \textit{Gaussian filter} is applied before the edge detection, obtaining a clear highlighting of the interesting edges (as seen in Fig.  \ref{fig:canny_gaussian} (C)).

@@

For this particular task, \textit{Keras}, an open-source Python library for Deep Learning \footnote{https://keras.io/}, was chosen to build the model. Many experiments were made to find the right architecture, most of them 
exploiting \textit{convolutional neural networks}. Above all of them, KODA uses \textit{U-Net}
\footnote{O. Ronneberger, P. Fischer, T. Brox, "U-Net: Convolutional Networks for Biomedical Image Segmentation", CoRR, abs/1505.04597, 2015}, a popular architecture for \textit{object segmentation}, and in particular an open-source implementation for Keras \footnote{https://github.com/zhixuhao/unet}.

@@

Due to the high number of required samples, the dataset itself is insufficient to fully train the network.
For this reason, a technique called \textit{data augmentation} is applied to provide the model
with enough variation. In particular, before being fed to the network, a random combination of transformations, such as scale, translation, rotation and brightness changes, is applied to the image and to the resulting edge feature map, as shown in Fig. \ref{fig:augmented_image}.

@@

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=\linewidth]{edge_comparison_epoch.png}
	\caption{Comparison of the edge detector over different epochs.}
	\label{fig:edge_comparison_epoch}
\end{figure}

@@

\begin{figure}[!htbp]
	\centering
	\makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{deep_edge_comparison.png}}%

@@

\section{Corner detection}

The corner detection task goal is to find the four document corners starting from the one-channel edges image. These points should be as precise as possible in order to define the tightest convex polygon around the document. The intuitive method to solve this task is to use the Hough transformation, as suggested in a Dropbox article \footnote{https://blogs.dropbox.com/tech/2016/08/fast-and-accurate-document-detection-for-scanning/}, to find straight lines that best approximate the document edges; the corners will be the intersection between those lines.

@@

To maintain uniformity for the client API, the corners are sorted following the Cartesian quadrants and they are scaled up based on the original image size provided as input.

@@

In order to implement the \textit{Optical Character Recognition} feature needed to highlight keywords, KODA uses \textit{Tesseract}\footnote{https://github.com/tesseract-ocr/tesseract}, an open-source OCR engine by Google. Moreover, it uses \textit{TesserOCR}\footnote{https://github.com/sirfz/tesserocr}, a Python wrapper which provides a fast binding to the C++ native library and exposes all the low-level APIs.
